MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-05-03 latex/paper/sections/08-netlist.tex
rw hello world

In FPGA context, the components are logical cells (\texttt{LUTs}, \texttt{CARRY4s}, etc.) or hierarchical cells (Verilog module instances) with pins connected together by wires. 
In Vivado, a Netlist can be synthesized as a Hierarchical or a Flattened netlist. 
Figure ~\ref{fig:hierarchical_design} shows an example a Verilog design using module instance hierarchy, while figure ~\ref{fig:hier_netlist} shows the same design synthesized into a flattened netlist. 
The synthesizer attempts to construct the module hierarchy as close to the module instantiation hierarchy as defined by the raw user design entry. 
Figure ~\ref{fig:flat_netlist} shows the same design but synthesized into a flat netlist. 
In either synthesized Netlist, the leaf cells, (deepest level cells), must necessarily consist only of primitive cells from the architecture primitives library (\texttt{LUT6}, \texttt{FDRE}, \texttt{CARRY4}, \texttt{DSP48E1}, etc.). 
The netlist can be compiled and exported as a purely structural low-level Verilog file, or an Electrinic Design Interchange Format (EDIF) file, both describing the netlist explicitly as a list of logical cells connected by a list of wires.

@@

\begin{lstlisting}[language=Java, caption={Netlist extraction}, label={lst:netlist_extract}]
Design design = Design.readCheckpoint("synth.dcp")
EDIFNetlist = design.getNetlist();
\end{lstlisting}

@@

SHOULD I INCLUDE A RAPIDWRIGHT HELLO WORLD EXAMPLE JUST TO GIVE AN IMPRESSION?

OR CAN I JUST DIRECT THE READER TO THE RAPIDWRIGHT DOCUMENTATION?

>>>>

In FPGA context, the components are logical cells (\texttt{LUTs}, \texttt{CARRY4s}, etc.) or hierarchical cells (Verilog module instances) with pins connected together by wires. 
In Vivado, a Netlist can be synthesized as a Hierarchical or a Flattened netlist. 
Figure ~\ref{fig:hierarchical_design} shows an example a Verilog design using module instance hierarchy, while figure ~\ref{fig:hier_netlist} shows the design synthesized into a hierarchical netlist with \textbf{hierarchical cells} and \textbf{leaf cells}. 
The synthesizer attempts to construct the module hierarchy as close to the module instantiation hierarchy defined by the user design entry. 
Figure ~\ref{fig:flat_netlist} shows the same design but synthesized into a flattened netlist.

In either synthesized netlist, the \textbf{leaf cells}, (deepest level cells), must necessarily consist only of \textbf{primitive cells} from the architecture's primitive cell library (\texttt{LUT6}, \texttt{FDRE}, \texttt{CARRY4}, \texttt{DSP48E1}, etc.). 
The netlist can be compiled and exported as a purely structural low-level Verilog file, or an Electrinic Design Interchange Format (EDIF) file, both describing the netlist explicitly as a list of logical cells connected by a list of wires.

@@

\begin{lstlisting}[language=Java, caption={Netlist extraction and traversal}, label={lst:netlist_extract}]
Design design = Design.readCheckpoint("synth.dcp")
EDIFNetlist = design.getNetlist();

// Example task:
// Extract the set of all nets from the design.

// Initialize a new Set:
Set<EDIFHierNet> netSet = new HashSet<>();

// Access every cell:
List<EDIFHierCellInst> ehcis = EDIFNetlist.getAllLeafHierCellInstances();

// Traverse cell list:
for (EDIFHierCellInst ehci : ehcis) {
    // Access every port on this cell:
    List<EDIFHierPortInst> ehpis = ehci.getHierPortInsts();
    for (EDIFHierPortInst ehpi : ehpis) {
        // Access the net on this port:
        EDIFHierNet ehn = ehpi.getHierarchicalNet();
        netSet.add(ehn);
    }
}

// Example downstream task:
// For each net, access the source cells and sink cells.
for (EDIFHierNet ehn : netSet) {
    // Access the source ports (there should only be one)
    List<EDIFHierPortInst> sourcePorts = ehn.getLeafHierPortInsts(true, false);
    // Access the source cell
    EDIFHierCellInst sourceCell = sourcePorts.get(0).getHierarchicalInsts();

    // Access the sink ports
    List<EDIFHierPortInst> sinkPorts = ehn.getLeafHierPortInsts(false, true);
    // Access the sink cells
    List<EDIFHierCellInst> sinkCells = new ArrayList<>();
    for (EDIFHierPortInst ehpi : sinkPorts) {
        sinkCells.add(ehpi.getHierarchicalInst());
    }
}

\end{lstlisting}

@@
